### 2018-09-04 practice-proofs/EH4.tex
Added files for practice proofs 4-9.

We explicitly construct a bijection mapping from the unit interval to the whole real line.

\begin{thm} \label{thm:BijectionRInterval}
	Define $f : (0,1) \to \RR$ by
	$$
		f(x) =\frac{1}{1-x} - \frac{1}{x}.
	$$
	The function $f$ is a bijection.
\end{thm}

\hintbox{
	If you can show that $f$ is strictly increasing, then that should give you injectivity.

	For surjectivity, you need to justify that, for every real number $y$, there is some $x\in(0,1)$ such that $f(x)=y$.
	The idea is to consider, for large $M>0$, $f(1/M)$ and $f(1-1/M)$ show that these are large positive and negative numbers, respectively.
	Then you can use the intermediate value theorem to conclude that because, for some $M$, $f(1/M)<y<f(1-1/M)$, the required $x$ must exist.
}