### 2018-09-04 practice-proofs/PracticeProofs9.tex

%---------DO NOT EDIT THIS INDENTED SECTION
	% Preamble
	\documentclass[11pt,reqno,oneside,a4paper]{article}
	\usepackage[a4paper,includeheadfoot,left=35mm,right=35mm,top=00mm,bottom=30mm,headheight=40mm]{geometry} %sets up the margins
	\input{../texHead-Proof-Standard} % Use the standard texHead for this module. You should not edit this file.
	\input{../texHead-Proof-Personal} % Use your personal additional macros. You may edit this file.
	%---The following code defines the title, author, and date of the document.
	\title{Practice proofs \#9}
	\author{Anonymous}
	\date{\today}   % Using \today automatically updates to the document's build date
%----------------------------------
%---------IF YOU WANT TO DEFINE YOUR OWN MACROS, YOU CAN DO SO EITHER IN ../texHead-Proof-Personal.tex OR FROM HERE ...

%---------... TO HERE
\begin{document}
\maketitle
\thispagestyle{fancy}

%-----------EDIT FROM HERE

\begin{abstract}
	\input{PracticeProofsStandardAbstract.tex}

	I fitted the whole model solutions onto 2 pages.
\end{abstract}

We argue that real multiplication, as defined through multiplication of rational Cauchy sequences, satisfies the corresponding five field axioms.

\begin{lem} \label{lem:ExistsNonzeroCauchySequence}
	Suppose that $x\in\RR\setminus\{0_\RR\}$.
	Then there is a rational $M>0$ and a Cauchy sequence $(x_n)_{n\in\NN}$ such every term has $|x_n|>M$ and $(x_n)_{n\in\NN}$ belongs to the equivalence class $x$.
\end{lem}

\begin{proof}
	%--------------EDIT THIS PART--------------%
	% Hint: The idea is to argue that there is a rational Cauchy sequence in $x$ for which at most finitely many terms are less than $M$.
	% Try contradiction; you should end up proving $x=0_\RR$.
	% After that, you can fix the finitely many ``problem'' terms without changing the equivalence class of your Cauchy sequence.
	%------------------------------------------%
\end{proof}

\begin{thm} \label{thm:RealMultiplication}
	Real multiplication, defined as the descent of multiplication of rational Cauchy sequences, obeys field axioms~\textup{(6)--(10)}.
\end{thm}

\begin{proof}
	%--------------EDIT THIS PART--------------%
	% Hint: Axioms~(6)--(9) essentially follow from the corresponding properties for rational numbers.
	% For axiom~(10), the tricky part is showing that the sequence you want to represent the multiplicative inverse is actually a Cauchy sequence.
	% It might be easiest to start by structuring the proof as if you have a magical lemma that guarantees the sequence is Cauchy, then go back and fill the gap, using lemma~\ref{lem:ExistsNonzeroCauchySequence} and some ingenuity.
	%------------------------------------------%
\end{proof}

\end{document}