### 2018-09-04 practice-proofs/PracticeProofsExtraHints.tex

%---------DO NOT EDIT THIS FILE
% Preamble
\documentclass[11pt,reqno,oneside,a4paper]{article}
\usepackage[a4paper,includeheadfoot,left=35mm,right=35mm,top=00mm,bottom=30mm,headheight=40mm]{geometry} %sets up the margins
\input{../texHead-Proof-Standard} % Use the standard texHead for this module. You should not edit this file.
\newcommand{\hintbox}[1]{\noindent\fbox{\begin{minipage}[t]{1.0\linewidth} \begin{hint}#1\end{hint} \end{minipage}}\clearpage}

\title{Extra hints for practice proofs}
\author{}
\date{\today}

\begin{document}
\maketitle
\thispagestyle{fancy}

This document will contain extra hints for all the practice proofs, beginning at practice proofs 4.
If it is ever out of date, please \href{mailto:\InstEmail ?subject=Please update 'Extra hints for practice proofs'}{email the instructor} to ask for an update.

\tableofcontents
\clearpage

\setcounter{section}{3}
\section{Practice proofs 4} \input{EH4} \setcounter{thm}{0}
\section{Practice proofs 5} \input{EH5} \setcounter{thm}{0}
\section{Practice proofs 6} \input{EH6} \setcounter{thm}{0}
\section{Practice proofs 7} \input{EH7} \setcounter{thm}{0} \clearpage
\section{Practice proofs 8} \input{EH8} \setcounter{thm}{0} \clearpage
\section{Practice proofs 9} \input{EH9} \setcounter{thm}{0} \clearpage

\end{document}