### 2018-09-04 practice-proofs/PracticeProofsStandardAbstract.tex

%---------DO NOT EDIT THIS FILE

The following exercise is meant to be challenging.
It is expected that your submission will not be perfect, but you will receive feedback on your submission so that you can improve it.
You \textbf{must} complete this exercise entirely on your own.
Do \textbf{not} collaborate with your peers on this exercise.
Do \textbf{not} ask the tutors for help, or search books / web for assistance.
There are hints in the .tex file.
There are more hints in the document \href{run:./PracticeProofsExtraHints.pdf}{PracticeProofsExtraHints.pdf}.
If you need help, please contact Dave by email: \href{mailto:\InstEmail}{\nolinkurl{\InstEmail}}.

You will be graded on participation.
That means it is OK to not have perfect answers, but you will get more out of the exercise if you try your best.

Be sure not to put your name anywhere on the document.
{\color{red}Before submitting, please delete the begin \& end abstract tags and everything between them.}