### 2018-10-03 practice-proofs/EH5.tex

\begin{prop} \label{prop:PowerSetsPreserveEquinumerosity}
	Suppose $\# X=\# Y$.
	Then $\# \power(X)=\# \power(Y)$.
	Moreover, if $\# X\leq\# Y$, then $\# \power(X)\leq\# \power(Y)$.
\end{prop}

\hintbox{
	There must be a bijection $f:X\to Y$.
	Pick an arbitrary $S\subset X$.
	Think about the process of taking the image of $S$ in $f$.
	Is that a function of $S$?
	Showing injectivity / surjectivity requires you to exploit injectivity / surjectivity of $f$.

	It is probably more efficient to prove the second part of the proposition first.
}

\begin{prop} \label{prop:SubsetCardinality}
	If $X\subset Y$, then $\# X\leq \# Y$.
\end{prop}

\hintbox{
	Directly construct an injection $f:X\to Y$.
	If the codomain of $f$ was $X$, then you could choose a bijection quite easily.
	What happens if you expand the codomain back to $Y$?
}

\begin{cor} \label{cor:IntersectionCardinality}
	$\# (X\cap Y) \leq \# X$.
\end{cor}

\hintbox{
	Use proposition~\ref{prop:SubsetCardinality}.
	How do subsets relate to intersections?
}